---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# rrdemoVersion1

<!-- badges: start -->
<!-- badges: end -->

# A reproducible research compendium example

This repository provides an example of a *reproducible research compendium*. The
analysis that it does is essentially a toy example, but it works and how it 
is produces is general enough (albeit with perhaps some small modifications and additions) 
to be useful for most small to medium sized data analysis projects.

One of the first papers to articulate the general concept of a reproducible
research compendium, especially one that uses R based tools, was [Statistical
Analyses and Reproducible
Research](https://biostats.bepress.com/bioconductor/paper2/) by Gentleman and
Temple Lang in 2004 (also published in 2007 in [Journal of Computational and
Graphical Statistics](https://www.jstor.org/stable/27594227). Gentleman and
Temple Lang defined a research compendium as an archive that includes all the
original raw data and computer code necessary to reproduce all the analyses
reported in a research study, but in addition includes one or more *dynamic
documents* that produce all of the reports (e.g., the manuscript for
publication, the supplementary documents, slides for presentations, etc) with
all the figures, tables, and other numerical or statistical quantities that
constitute the communicated results of the analysis. From a compendium, anyone
ought to be able to exactly reproduce every step of the analysis, including
intermediate results and data sets, on the basis of the original raw data, and
also should be able to determine exactly how any figure, table, numerical and
statistical etc any reports was obtained.

When Gentleman and Temple Lang wrote [Statistical Analyses and Reproducible
Research](https://biostats.bepress.com/bioconductor/paper2/) many tools that
facilitate the creation of a research compendium were either not well known or
simply did not exist at all. These tools include *Git*, *GitHub*, *RMarkdown*,
*knitr*, *pandoc*, *devtools*, *docker*, *travis*, etc. As a result, at that
time, producing a research compendium was a time consuming and arduous task. By
contrast, now that the tools just mentioned are well known and widely used,
producing a research compendium is remarkably easier. Of course, it still
requires effort and commitment, but this is well worth it because of the
transparency and quality control guarantees that it brings to your work and
because it will allow others, including and especially your future self, to
understand and reproduce your analysis exactly.


# How to use this compendium

There are a number of different ways to use this compendium. In any case, the
first step is always the same: clone the repository from GitHub.
``` 
git clone https://github.com/mark-andrews/rrcompendium.git
```
Then open the *RStudio project* that is within the compendium by going to *File /
Open Project* in RStudio and choosing the `rrcompendium.Rproj` file.

## Use the `analysis.Rmd` RMarkdown file in an normal RStudio session

The simplest way of using this compendium is to just open the `analysis.Rmd`
file in a normal RStudio session, and either render to a pdf or use it
interactively as a
[notebook](https://bookdown.org/yihui/rmarkdown/notebook.html). Proceeding this
way could work just fine. However, even under normal circumstances it will
requires some manual installation of R packages, as we'll see. But more,
when things start getting complex --- when there are lots dependencies, or when
very specific versions of dependencies are required, or when additional external
software or libraries are required --- this simple way might not work, or at
least without a lot of time and effort and patience on the part of the user.

To proceed in this normal ways. 
You can open the *RStudio project* in the compendium by going to *File /
Open Project* in RStudio and choosing the `rrcompendium.Rproj` file. 

You can
then open the `analysis.Rmd` RMarkdown file that is in the `analysis` directory. 

To be able to use `analysis.Rmd` fully, you first need to ensure your your R/RStudio is set
up rendering RMarkdown documents to pdf. You require the R package
`rmarkdown`, but also you need the external software [pandoc](https://pandoc.org/) and
[TeX/LaTeX](https://www.latex-project.org/). RStudio (since version v0.98.932)
comes bundled with `rmarkdown` and `pandoc`, so you may have these already. 
However, it is generally a good idea to use
the latest version of RStudio always and always keep your R packages up to date.
You can do the latter from the RStudio console as follows:
```{r, eval=FALSE}
update.packages(ask = FALSE)
```
Or, to just update `rmarkdown`, do 
```{r, eval=FALSE}
install.packages('rmarkdown')
```
Installing [TeX/LaTeX](https://www.latex-project.org/) can be done in different 
ways depending on the platform (Windows, Mac OSX, Linux) that you use. 
If, however, 
you just want to use TeX/LaTeX within RStudio for rendering RMarkdown files
to pdf, then (by far) the easiest option is to install [tinytex](https://yihui.name/tinytex/) as follows:
```{r, eval=FALSE}
install.packages('tinytex')
tinytex::install_tinytex()
```

You also need to install the special purpose `rrcompendium` R package that
accompanies this compendium. This is installed as follows:
```{r, eval=FALSE}
devtools::install_github('mark-andrews/rrcompendium')
```
Installing this package will also install all the R dependencies you need. 

Now you can render `analysis.Rmd` to a pdf by pressing the *Knit* button on the
source file editor tab, or select `File / Knit Document`, or (best option, in my
opinion) press `Ctrl + Shift + K`.

You can also 
